---
title: "Srtoop deadline: creating processed data from raw"
author: "Zsuzsa Szekely"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
editor_options: 
  chunk_output_type: console
---

# Load packages

```{r load packages}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, osfr, janitor, lubridate, purrrlyr)
```

# Load helper functions

```{r load helper functions}
source("R/utils.R")
```

# Import data

```{r}
# Reading in data files
raw <- read_csv("data/deadline/raw/raw_data.csv")
```

# Data sorting

## make only trial response data stay in the dataframe, remove all other rows (all the trials have a value in variable "word", all the other rows have no value there)

```{r}
response <- 
  raw %>% 
  filter(!is.na(word))
```

## create variable for identifying the different task sections (practice, calibration, test)

```{r}
response <- 
  response %>% 
  mutate(section = case_when(sender == "StroopPracticetrial" ~ "practice",
                             sender == "1_StroopCalibrationTrial" ~ "calibration",
                             sender == "2_StroopCalibrationTrial" ~ "calibration",
                             sender == "3_StroopCalibrationTrial" ~ "calibration",
                             sender == "4_StroopCalibrationTrial" ~ "calibration",
                             sender == "1_StroopCalibrationTrial" ~ "calibration",
                             sender == "1_StroopTestTrial" ~ "test",
                             sender == "2_StroopTestTrial" ~ "test",
                             sender == "3_StroopTestTrial" ~ "test",
                             sender == "4_StroopTestTrial" ~ "test"
                             ))
```

# create variable for trial serial number within each section of each block

```{r}
response <- 
  response %>% 
  group_by(participant_id, blockId, section) %>%
  mutate(trialNumber = row_number()) %>% 
  ungroup()
```

# add variable storing whether the previous trial was congruent

```{r previous congruency}
response <-
  response %>%
  mutate(isPrevCongruent = case_when(trialNumber < 2 ~ NA_integer_,
                                     trialNumber >= 2 & lag(congruency) == "con" ~ 1L,
                                     trialNumber >= 2 & lag(congruency) == "inc" ~ 0L))
```

# Create condition variable based on the congruency of current and previous trials

* ii: incongruent follows incongruent
* ic: congruent follows incongruent
* ci: incongruent follows congruent
* cc: congruent follows congruent

```{r}
response <-
  response %>%
  mutate(condition = case_when(isPrevCongruent == 0L & congruency == "inc" ~ "ii",
                               isPrevCongruent == 0L & congruency == "con" ~ "ic",
                               isPrevCongruent == 1L & congruency == "inc" ~ "ci",
                               isPrevCongruent == 1L & congruency == "con" ~ "cc",
                               TRUE ~ NA_character_))
```

# make timeout (NA) responses FALSE in variable "correct", so that timeout responses count as incorrect responses

```{r}
# we expect to have missing data in "correct" variable only in case of timeout trials (that has no response). check whether there's any missing data in case of non-timeout trials
response %>%
  count(is.na(correct))

response %>%
  count(is.na(response))

# replace NA to FALSE in variable "correct"
response <- 
  response %>%
  mutate(correct = replace_na(correct, FALSE))

# check if there's any NA-s remaining in the response variable (all responses should have the value of TRUE or FALSE)
response %>% 
  count(is.na(correct))

response %>% 
  count(correct)
```

# create new variable storing whether the previous trial was correct

```{r}
response <- 
  response %>% 
  mutate(isPrevCorrect = case_when(trialNumber < 2 ~ NA_integer_,
                                   trialNumber >= 2 & lag(correct) == TRUE ~ 1L,
                                   trialNumber >= 2 & lag(correct) == FALSE ~ 0L))
```








PARTICIPANT LEVEL EXCLUSIONS

* 1. participants who didn't finish the task will be excluded

(This step was conducted in the source_raw file, now we just check it.)
Each participant who completed the task should have data from 24 practice trials, 112 calibration trials and 324 test trials (total = 460).

# check if the task was completed by everyone

```{r}
response %>% 
  group_by(participant_id) %>% 
  count(section) %>%
  mutate(complete = case_when(section == "practice" & n == 24 ~ "complete",
                              section == "calibration" & n == 112 ~ "complete",
                              section == "test" & n == 324 ~ "complete", 
                              TRUE ~ "incomplete"
                              )) %>% 
  ungroup() %>% 
  count(complete) # the result should equal to the number of participants (57) * 3
```

* 2. participants with mean accuracy below 70% of the mean accuracy of the whole sample will be excluded

# calculate accuracy of respondents and create variable (dropParticipantAcc) for accuracy-based exclusion. accuracy is calculated in case of calibration and test trials (therefore we use cal_test df)

```{r}
response %>%
  group_by(participant_id, correct) %>%
  summarise(n = n()) %>%
  ungroup() %>% 
  group_by(participant_id) %>%
  mutate(N = sum(n),
         prop = n / N * 100,
         dropParticipantAcc = case_when(prop < 70 ~ 1L,
                                        TRUE ~ 0L))

response <- 
  response %>% 
  group_by(participant_id) %>%
  mutate(prop = sum(correct) / n() * 100,
         dropParticipantAcc = case_when(prop < 70 ~ 1L,
                                        TRUE ~ 0L)) %>% 
  ungroup()
```

# Number of participants dropped because of accuracy.

```{r}
response %>% 
  distinct(participant_id, dropParticipantAcc) %>% 
  group_by(dropParticipantAcc) %>%
  summarise(n = n()) %>% 
  ungroup()
```

# Get the id of the participants who will be dropped.

```{r}
response %>%
  filter(dropParticipantAcc == 1) %>%
  distinct(participant_id) # id of participants to be dropped: 7, 16, 42
```

# Exclusion

```{r}
response <-
  response %>%
  group_by(participant_id) %>% 
  filter(!any(dropParticipantAcc == 1L))
```

Check if the exclusion was succesful. There should be 54 respondents.
  
```{r}
response %>% 
  distinct(participant_id) %>% 
  nrow()
```

* 3. participants with mean RT more than 2.5 SDs away from the mean RT of the whole sample will be excluded

We use the RT-s of the the calibration trials only, as their mean RT give the personal deadline for the test trials.

First, we check if personal deadlines equal to the mean RT-s of the matching calibration sections

```{r}
response %>% 
  group_by(participant_id, section, blockId) %>%
  mutate(mean_duration = mean(duration, na.rm = TRUE)) %>%
  distinct(participant_id, section, blockId, duration, personalDeadline, mean_duration) %>% 
  ungroup() %>% 
  View()
```

Calucalte the mean response time and SD for exclusion

We calculate the mean RT for each participant and the grand mean RT for participant level exclusion.

```{r calculate mean response time}
participant_rt <- response %>%
  group_by(participant_id) %>%
  filter(section == "calibration") %>% 
  mutate(personal_mean_duration = mean(duration, na.rm = TRUE),
         personal_sd_duration = sd(duration, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(grand_mean_duration = mean(duration, na.rm = TRUE),
         grand_sd_duration = sd(duration, na.rm = TRUE))

# create variable indicating outliers
participant_rt <- participant_rt %>% 
  group_by(participant_id) %>% 
  mutate(drop_participant_rt = case_when(grand_mean_duration + 2.5 * grand_sd_duration < personal_mean_duration ~ 1L,
                                         grand_mean_duration - 2.5 * grand_sd_duration > personal_mean_duration ~ 1L,
                                         TRUE ~ 0L)) %>%
  ungroup()
```

Get the number of rt outliers

```{r}
participant_rt %>% 
  filter(drop_participant_rt == 1) %>%
  nrow()
```

Since there is nobody to be excluded, the next 4 steps are not required.

Get the number of participants per task before exclusion.

```{r}
# participant_rt %>%
#   distinct(participant_id) %>%
#   nrow()
```

Get the id of the participants who will be dropped.

```{r}
# participant_rt %>%
#   filter(drop_participant_rt == 1) %>%
#   distinct(participant_id)
```

Exclusion.

```{r participant level RT exclusion}
# response <- response %>% 
#   group_by(participant_id) %>% 
#   filter(!participant_id == "")
```

Check if the exclusion was successful. There should be ... participants remaining.
  
```{r}
# response %>% 
#   distinct(participant_id) %>% 
#   nrow()
```








TRIAL LEVEL EXCLUSIONS

* 1. exclude practice and calibration trials
* 2. exclude first response trial in each block for each participant
* 3. exclude rt outlier trials
* 4. exclude trials directly following outlier trials

For the rt analysis there are some more exclusion criteria:
* 5. exclude error trials
* 6. exclude trials directly following error trials


* 1. exclude practice and calibration trials

Save practice, calibration and test trials to separate dfs

```{r}
practice <-
  response %>%
  filter(section == "practice")

calibration <-
  response %>%
  filter(section == "calibration")

test <-
  response %>%
  filter(section == "test")

# df containing both calibration and test trials
cal_test <- 
  response %>% 
  filter(section != "practice")
```

The number of trials after separating trials:

Expected number of trials for each participant:
 * practice: 24
 * calibration: 112
 * test: 324

```{r}
# practice %>%
#   group_by(participant_id) %>%
#   count() %>%
#   View()
# 
# # practice %>%
# #   group_by(participant_id) %>%
# #   summarise(n = nrow(.)) %>%
# #   View()
# 
# calibration %>%
#   group_by(participant_id) %>%
#   count() %>%
#   View()
# 
# test %>%
#   group_by(participant_id) %>%
#   count() %>%
#   View()
# 
# cal_test %>%
#   group_by(participant_id) %>%
#   count() %>%
#   View()
```

* 2: exclude first test trial in each block for each participant

The task had 4 blocks. Therefore, the number of trials that will be dropped equals the number of participants multiplied by four. Lets check this.

```{r}
# get the number of participants
participant_rt %>%
  distinct(participant_id) %>%
  nrow()

# get the number of trials
nrow(test)

# get the number of first trials
test %>%
  filter(trialNumber == 1) %>%
  nrow()
```

Dropping first trials.

```{r}
test <- 
  test %>% 
  filter(trialNumber != 1) # 1st trial level exclusion criteria

# Other way to do this: !is.na(isPrevCongruent)

# check the number of first trials (it should be 0)
test %>%
  filter(trialNumber == 1) %>%
  nrow()

# check the number of remaining trials
nrow(test)
```

* 3: prepare exclusion of trials with RT more than 2.5 SDs away from conditional mean RT

Calculate the mean RT per condition (iI, iC, cI, cC)

```{r}
test <-
  test %>%
  group_by(participant_id, condition) %>% 
  mutate(rtConditionalMean = mean(reaction_time, na.rm = T),
         rtConditionalSd = sd(reaction_time, na.rm = T)) %>% 
  ungroup()
```

Create variable that stores whether a participant or a trial should be dropped based on the rt criteria

```{r}
test <- test %>% 
  group_by(participant_id) %>%
  mutate(dropTrialRt = case_when(rtConditionalMean + 2.5 * rtConditionalSd < reaction_time ~ 1L,
                                 rtConditionalMean - 2.5 * rtConditionalSd > reaction_time ~ 1L,
                                 TRUE ~ 0L),
# At this point there should be no trial == 1 trials because they were removed, trial == 2 trials will have NO previous trial as the immediately preceding trial will be the last trial of the previous block
         isPrevDropTrialRt = case_when(trialNumber == 2 ~ NA_integer_, 
                                       trialNumber > 2 & lag(dropTrialRt) == 1L ~ 1L,
                                       trialNumber > 2 & lag(dropTrialRt) == 0L ~ 0L)) %>% 
  ungroup()
```

Lets take a look at the number of the congruent and incongruent trials per condition.

```{r}
test %>% 
  group_by(condition) %>% 
  count()
```

* Drop outliers (trials 2.5 SDs away from the conditional mean of the participant)

```{r}
test <-
  test %>% 
  filter(dropTrialRt == 0L)
```

* 4. Drop trials following outlier trials

```{r trial level exclusion reaction time}
test <-
  test %>% 
  filter(isPrevDropTrialRt == 0L)
```

Check the number of the congruent and incongruent trials per condition again

```{r}
test %>% 
  group_by(condition) %>% 
  count()
```



# acc analysis exclusions

For the accuracy analysis there's no more exclusion criteria. We save the data into a separate df for accuracy analysis and we check if after the exclusions there is still some variance in the accuracy scores.

```{r}
testAccAnalysis <- test

testAccAnalysis %>% 
  group_by(condition) %>% 
  summarise(mean_acc = mean(correct, na.rm = TRUE),
            sd_acc = sd(correct, na.rm = TRUE))

testAccAnalysis %>% 
  group_by(congruency) %>% 
  summarise(mean_acc = mean(correct, na.rm = TRUE),
            sd_acc = sd(correct, na.rm = TRUE))
```

## Checking whether there are trials remaining in each condition (ii, ic, ci, cc) from each participant for the accuracy analysis

```{r}
testAccAnalysis %>% 
  count(participant_id, condition) %>% # get the number of trials per each participant in each condition
  filter(n != 0) %>% # drop those conditions with 0 trials
  group_by(participant_id) %>%
  summarise(n = n()) %>%  # get the number of remaining conditions per each participant
  filter(n != 4) %>% # see if there's anybody with less than 4 conditions
  nrow() # get the number of these participants
```

# Looking for ceiling effect regarding accuracy
## Mean and SD of accuracy of congruent and incongruent trials

```{r}
# creating variable (congruency_prop) for accuracy measured in congruent and incongruent trials
testAccAnalysis <-
  testAccAnalysis %>% 
  group_by(participant_id, congruency) %>% 
  mutate(congruency_prop = sum(correct) / n()*100,
         congruency_sd = sd(correct))

# calculating mean and SD of accuracy of congruent and incongruent trials 
testAccAnalysis %>% 
  group_by(congruency) %>% 
  summarise(mean_acc = mean(congruency_prop),
            sd_acc = sd(congruency_prop),
            n_trial = n())
```

## Plots of accuracy distribution

```{r}
testAccAnalysis %>% 
  group_by(participant_id) %>% 
  ggplot(aes(x = prop)) +
  geom_bar() +
  scale_y_continuous(limits = c(0, 1500)) +
  scale_x_continuous(limits = c(70, 100))
```



# rt analysis exclusions

For the rt analysis there are some more trial level exclusion criteria.

Create df for rt analysis, then exclude:
* 5. Error trials

```{r}
testRtAnalysis <-
  test %>% 
  filter(correct == TRUE)
```

* 6. Trials following error trials

```{r}
testRtAnalysis <-
  testRtAnalysis %>% 
  filter(isPrevCorrect == 1L)
```

## Checking whether there are trials remaining in each condition (ii, ic, ci, cc) from each participant for the rt analysis

```{r}
testRtAnalysis %>% 
  count(participant_id, condition) %>% # get the number of trials per each participant in each condition
  filter(n != 0) %>% # drop those conditions with 0 trials
  group_by(participant_id) %>%
  summarise(n = n()) %>%  # get the number of remaining conditions per participants
  filter(n != 4) %>% # see if there's anybody with less than 4 conditions
  nrow() # get the number of these participants
```








Save rt and acc analysis to separate dataframes

```{r save tasks}
# Saving data for the reaction time analysis
write_tsv(testRtAnalysis, "data/deadline/processed/processed_data_rt.tsv")

# Saving data for the accuracy analysis
write_tsv(testAccAnalysis, "data/deadline/processed/processed_data_acc.tsv")
```
